---
title: "Within-person variation analyses for thesis"
author: "R-Walmsley"
date: "29/03/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Note location of input files
```{r}
file_loc <- "../../../Repeated_measurement_error_related/repeatedMeasurements/R/"
file_loc_local <- "../../useful_functions/"
input_loc <- "../../../../paperRW2021/"
input_loc_rm <- "../../../Repeated_measurement_error_related/repeatedMeasurementsProcessing/data/59070/"
```


# Load packages
```{r echo = FALSE, message = FALSE, warning = FALSE}
library(survival)
library(ggplot2)
library(data.table)
library(table1)
library(knitr)
```

# Prepare functions for use in adjustment 
```{r message=FALSE, results='hide'}
fl <- list.files(file_loc)
for (file in fl){
  try(source(paste0(file_loc, file)))
}
# Sourcing functions from another project - to be replaced with install_github
```

# Load helpers
```{r results = 'hide'}
fl <- list.files(file_loc_local)
for (file in fl){
  try(source(paste0(file_loc_local, file)))
}
```

# Read in data 
Read in main data:
```{r}
# Prepped UKB data
df <- readRDS(paste0(input_loc, "epiAnalysis/inputData/2021-06-17check_messages_ready_to_use.RDS"))
```

Read in repeated measurements sample:
```{r}
df_rm <-
  data.frame(data.table::fread(paste0(input_loc_rm, "summary_data_all_with_5sec_epoch_and_ml_oct21.csv")))
```

# Quality at each repeat
```{r}
for (i in 0:4){
  good_data_cond <- (!(is.na(df_rm[, paste0("acc.overall.avg_s", i)]))) &
                  (df_rm[, paste0("quality.goodCalibration_s", i)] == 1) & 
                  (df_rm[, paste0("quality.goodWearTime_s", i)] == 1) & 
                  (df_rm[, paste0("clipsBeforeCalibration_s", i)] < 0.01*df_rm[, paste0("totalReads_s", i)]) & 
                  (df_rm[, paste0("clipsAfterCalibration_s", i)] < 0.01*df_rm[, paste0("totalReads_s", i)]) & 
                  (df_rm[, paste0("acc.overall.avg_s", i)] <100)
  df_rm[, paste0("in_s", i)] <- ifelse(good_data_cond, TRUE, FALSE)
}
df_rm$any_valid_repeat <- df_rm$in_s1 |
                                 df_rm$in_s2 | 
                                 df_rm$in_s3 | 
                                 df_rm$in_s4
```

# Exclusions in repeated measures data
```{r}
n <- nrow(df_rm)
exc_rm <-
  data.frame("Reason" = "Starting Number",
             "n excluded" = NA,
             "n remaining" = n)

#--------------------
df_rm <- df_rm[is.na(df_rm$file.startTime_s1)|(df_rm$file.startTime_s1 >= as.Date("2017-01-01")), ]
exc_rm <- rbind(exc_rm, 
  data.frame("Reason" = "s1 file appears erroneous",
             "n excluded" = n - nrow(df_rm),
             "n remaining" = nrow(df_rm)))
n <- nrow(df_rm)

#---------------------
df_rm <- df_rm[df_rm$in_s0, ]
exc_rm <- rbind(exc_rm, 
  data.frame("Reason" = "Missing baseline",
             "n excluded" = n - nrow(df_rm),
             "n remaining" = nrow(df_rm)))
n <- nrow(df_rm)

#---------------------
df_rm <- df_rm[df_rm$any_valid_repeat,]
exc_rm <- rbind(exc_rm, data.frame("Reason" = "Missing/poor quality data for all repeats",
             "n excluded" = n - nrow(df_rm),
             "n remaining" = nrow(df_rm)))
n <- nrow(df_rm)


#---------------------
df_rm <- df_rm[df_rm$eid %in% df$eid,]
exc_rm <- rbind(exc_rm, 
  data.frame("Reason" = "Not in final analytic sample",
             "n excluded" = n - nrow(df_rm),
             "n remaining" = nrow(df_rm)))


#---------------------
print(exc_rm)
```


# Repeated measurements data prep
- Add ilr transformed variables:
```{r}
# List of elements -----------------------------------------------------------------------------------------------------------
rot_list <- list(
  "sleep" = c("sleep", "sedentary", "light", "MVPA"),
  "sedentary" = c("sedentary", "sleep", "light", "MVPA"),
  "light" = c("light", "sleep", "sedentary",  "MVPA"),
  "MVPA" = c("MVPA", "sleep", "sedentary", "light")
) 

# The class labels in the repeated measurements dataset are different to in the main dataset, so relabel the main dataset to match the repeated measurements dataset ---------------------------------------------------------------------------------------------------------
df$sedentary <- df$SB
df$light <- df$LIPA
for (var in rot_list[[1]]){
  df[, paste0(var, "_ML4_red_s0")] <- df[, var]
}


# Cycle through different first pivot coordinates adding in the set of pivot coordinates with this first pivot coordinate----------------------
for (j in 1:length(rot_list)){
  for (i in 0:4){
    # Add in transformed variables for each repeat----------------------
    comp_labels <- paste0(rot_list[[j]], "_ML4_red_s", i)
    tl <- epicoda::transf_labels(comp_labels, transformation_type = "ilr")
  
    # Det limit for zero imputation------------------------------------
    dl <- min(df$MVPA[df$MVPA>0]) # This is coming from the main study as a better overall source
    
    # Set up a mask for rows which will not appear----------------------
    df_rm[, paste0(tl, i)] <- NA
    inc_ind <- !(df_rm[, paste0("excluded_s", i)]) & !(is.na(df_rm[, paste0("excluded_s", i)]))

    # Do and store transformation---------------------------------------
    d <- epicoda::transform_comp(df_rm[inc_ind, ], comp_labels = comp_labels, det_limit = dl)[, tl]

    # Put these back in the data frame with a label for the repeat-----
    df_rm[inc_ind, paste0(tl, "_s", i)] <-  d
    df_rm[, paste0(names(rot_list)[j], c("z1", "z2", "z3"), "_s", i)] <- df_rm[, paste0(tl, "_s", i)]
    
    # Only on the first repeat, also transform the data in the main dataset-------------
    if (i == 0){
      d2 <- epicoda::transform_comp(df, comp_labels = comp_labels, det_limit = dl)[, tl]
      df[, paste0(names(rot_list)[j], c("z1", "z2", "z3"), "_s", i)] <- d2[, tl]
    }

  }
}

# Note that the labelling of columns is a little weird - MVPAz2_s0 refers to z2 for the case where MVPA was the pivot at repeat 0
```

# Descriptive table for repeated measurements data 
- Preprocessing
```{r}
# FORMAT AS DATE WHERE NEEDED
for (i in 0:4) {
  df_rm[, paste0("file_start_s", i)] <-
    as.Date(df_rm[, paste0("file.startTime_s", i)])
}

# MERGE IN SOME DATA
nrow(df_rm)
df_rm <- merge(df_rm, df[, c("eid", colnames(df)[!(colnames(df) %in% colnames(df_rm))])], by = "eid", sort = FALSE)
nrow(df_rm)

# SET UP VARS TO PROCESS
vars_list <- list(
  "Wear date" = "file_start_s",
  "Overall average acceleration (mg)" = "acc.overall.avg_s",
  "ML moderate-to-vigorous physical activity (hr/day)" = "MVPA_ML4_red_s",
  "ML light physical activity (hr/day)" = "light_ML4_red_s" ,
  "ML sedentary behaviour (hr/day)" = "sedentary_ML4_red_s",
  "ML sleep (hr/day)" = "sleep_ML4_red_s",
  "First pivot coordinate: MVPA vs rest" = "MVPAz1_s",
  "First pivot coordinate: LIPA vs rest" = "lightz1_s",
  "First pivot coordinate: SB vs rest" = "sedentaryz1_s",
  "First pivot coordinate: sleep vs rest" = "sleepz1_s"
  
)
```


- 
```{r}
# Regression calibration procedure
## Regression calibration in full dataset
for (var in names(rot_list)) {
  # Find names of variables
  form_vars <-
    colnames(df_rm)[grepl(paste0(var, "z"), colnames(df_rm))]
  form_vars <- unique(gsub("_(.*)", "", form_vars))
  
  # Unadulterated Cox model
  form_vars_old <- paste0(form_vars, "_s0")
  form_sum_old <-  paste0(form_vars_old, collapse = "+")
  cox_basic <-
    coxph(formula(
      paste0(
        "Surv(age_entry, age_exit, CVD_event) ~ ",
        form_sum_old,
        "+ strata(sex) +
  ethnicity + smoking + alcohol + fruit_and_veg_cats + oily_fish + red_and_processed_meat_cats + education_cats + TDI_quartiles"
      )
    ), data = df)
  
  # Manually invert regression calibration using the matrix
  # Make models
  models <- calc_models_multi_single(
    var_names = form_vars,
    data = df_rm,
    suffix_baseline = "_s0",
    suffix_rep = c("_s1", "_s2", "_s3", "_s4"),
    exclusion_indic = "excluded",
    covariates = c("age_entry", "sex", "ethnicity", "smoking", "alcohol", "fruit_and_veg_cats", "oily_fish", "red_and_processed_meat_cats", "education_cats", "TDI_quartiles")
  )
  
  mat <- models$coefficients
  inverse_mat_relevant <- solve(mat[25:27, ])
  coef_before <- summary(cox_basic)$coefficients[1:3, 1]
  changed_coef <- inverse_mat_relevant %*% coef_before
  print(var)
  print(coef_before)
  print(changed_coef)
  
}
```

These results are nearly, but not quite, identical to what you get doing it the substitute-in way. Similar enough I'm happy they're correct. 

Why aren't they exact? The issue is because we adjust for age by using it as the timescale in the model. If we don't do that, we get identical: 

```{r}
# Regression calibration procedure
## Regression calibration in full dataset
for (var in names(rot_list)) {
  # Find names of variables
  form_vars <-
    colnames(df_rm)[grepl(paste0(var, "z"), colnames(df_rm))]
  form_vars <- unique(gsub("_(.*)", "", form_vars))
  
  # Unadulterated Cox model
  form_vars_old <- paste0(form_vars, "_s0")
  form_sum_old <-  paste0(form_vars_old, collapse = "+")
  cox_basic <-
    coxph(formula(
      paste0(
        "Surv(age_entry, age_exit, CVD_event) ~ ",
        form_sum_old, "+ age_entry + strata(sex) +
  ethnicity + smoking + alcohol + fruit_and_veg_cats + oily_fish + red_and_processed_meat_cats + education_cats + TDI_quartiles"
      )
    ), data = df)
  
    # Cox model with regression calibration
   # Calculate regression calibration model 
    models_list <- calc_models_multi(
    var_names = form_vars,
    data = df_rm,
    suffix_baseline = "_s0",
    suffix_rep = c("_s1", "_s2", "_s3", "_s4"),
    exclusion_indic = "excluded", 
    covariates = c("age_entry", "sex", "ethnicity", "smoking", "alcohol", "fruit_and_veg_cats", "oily_fish", "red_and_processed_meat_cats", "education_cats", "TDI_quartiles"))
  
  # Make predictions
  pred_data <-
    pred_rep(models_list, newdata = df, suffix_pred = "_pred")
  
  
  # Variables for new calculation
  form_vars_new <- paste0(form_vars, "_pred")
  form_sum <- paste0(form_vars_new, collapse = "+")
  
  cox_repmeas <-
    coxph(formula(
      paste0(
        "Surv(age_entry, age_exit, CVD_event) ~ ",
        form_sum, "+ age_entry + strata(sex) +
  ethnicity + smoking + alcohol + fruit_and_veg_cats + oily_fish + red_and_processed_meat_cats + education_cats + TDI_quartiles")
    ), data = pred_data)
   
  
  
  # Manual calculation
  # Make models
  models <- calc_models_multi_single(
    var_names = form_vars,
    data = df_rm,
    suffix_baseline = "_s0",
    suffix_rep = c("_s1", "_s2", "_s3", "_s4"),
    exclusion_indic = "excluded", 
    covariates =  c("age_entry", "sex", "ethnicity", "smoking", "alcohol", "fruit_and_veg_cats", "oily_fish", "red_and_processed_meat_cats", "education_cats", "TDI_quartiles")
  )
  
  mat <- models$coefficients
  inverse_mat_relevant <- solve(mat[25:27, ])
  coef_before <- summary(cox_basic)$coefficients
  changed_coef <- inverse_mat_relevant %*% coef_before[1:3]
  
  # Compare results
  print(var)
  print(changed_coef)
  print(summary(cox_repmeas)$coefficients)
}
```
See green notebook p111 - p112 for proof of why this is so. I think the way it's done is defensible in light of (a) aiming for consistency with earlier analyses (i.e. adjusting for same stuff in main Cox model) and (b) the fact that in regression calibration you can adjust for predictors of X that are not predictors of Y (see Keogh et al, STRATOS document, p24). 